Making multiple-locus association mapping on a genome-wide scale routine

### 2016-09-02 AM+draft.tex
02/09/16 Its been 6 months since I looked at this paper. Comitting prior to making major changes to paper

\subsection{Computational performance}

To assess the computational performance of am+, we conducted a simulation study. Our performance measures are computing time and memory usage. We also compared the performance of am+ to six other computer programs for multiple locus association mapping (table X).

Data were generated under five different scenarios. These scenarios are motivated by the type and size of gwas currently being performed. The five scenarios are  unsequenced crop plants (UCP, 150 plants and 5k snp), model plants ( 4555 plants and 34544k snp), ..... Our marker data are from the 1000 Genomes project (Phase 3). Quantitative trait data are generated from these data by assigning effect sizes to a randomly selected set of snp loci along with a gaussian deviate for the error. 
The size of the error is  set to yield a heritability of 0.5.

>>>>

\subsection{Computational performance}
We conducted a simulation study to assess  the computational performance of am+. For comparison, we assessed six other computer programs (table X) that  also implement multiple locus association mapping. The marker data in our simulation study comes from phase 3 of the 1000 genomes project. Here, 2315 individuals were sequenced and their genotype information made available. We generated trait data by randomly selecting snp, assigning allelic  effects to the selected snp, and aggregating these effects along with random error. The number of snp  selected  was sampled from a Poisson distribution with mean 30. The effect sizes were sampled from a beta distribution. The errors were drawn from a normal distribution with zero mean and whose variance yielded a quantitative trait with a heritability of 0.5.

We were interested in how study size impacts performance. We generated replicates under five different study sizes. These were 150 individuals and 4000 snp (150 X 4K), 300 individuals and 50000 snp (300 X 50k), ... We based these five study sizes on the size of study being performed in unsequenced crop plants such as wheat, ......, respectively. For those study sizes that exceeded the number of genotyped individuals in the 1000 genomes project,  we simulated additional data by forming genotype mosaics from the real snp data. We generated 250 replicates for each scenario.

Performance was measured by recording elapse  time and maximum memory usage. Analyses were performed on a high end desk top which was a dell intra quad core processor with 120G of RAM and a Intel gpu. Programs were run for a maximum of 100 hours.  All programs except piMass  made use of distributed computing, either implicitly via multithreaded blas and lapac linear algebra libraries or as in the case of ranfor, explicitly. For ranfor, we wrote R code to parallelise the random forest analysis across multiple cpu.  All scripts for performing these analyses are available from the Csiro-data acces portal.

In figure X we compare the elapse time of  am+ against the other computer programs. On the y axis, we have the ratio of each computer program's elapse time to the elapse time of cpu-based am+. On the X axis, we have the number of genotypes and study size. Both axes are on log scale . A positive (negative) log ratio means am+ had a shorter (longer) elapse time  than the other computer program. The magnitude (value) of the ratio is (the magnitude of)  how many times shorter or longer am+ is to the other computer program.  Since performance varies with replicate, for a given study size, a point in figure X is the log median of 250  ratios of elapse times.

From figure X, the superior performance of am+ to the other computer programs is evident. Describe ...
